---
title: "EDA Notebook"
output: html_notebook
---


```{r}
# loading tables:  assay_component_endpoint table from invitrodb
# & enrichment data (from Ryan Lougee[unreproducible])
library(data.table)
# queried from invitrodb 2020/12/18 14:00 PM
# read assay_component_endpoint table and its col names
aeid_table <- fread('InputData/aeid_table.txt')
aeid_table_col <- fread('InputData/aeid_table_col.txt')
# pasting the content instead of cleaning up aeid_table_col
col_names <- c('aeid', 'acid', 'assay_component_endpoint_name', 'export_ready', 'internal_ready', 'assay_component_endpoint_desc', 'assay_function_type', 'normalized_data_type', 'analysis_direction', 'burst_assay', 'key_positive_control', 'signal_direction', 'intended_target_type', 'intended_target_type_sub', 'intended_target_family', 'intended_target_family_sub', 'fit_all', 'cell_viability_assay', 'data_usability')
names(aeid_table) <-col_names

# enrichment data
baseline <- fread("InputData/ENRICHMENTS_invitrodb_v3_1_BASELINE_hitc_MC_2019.tsv")
burst <- fread("InputData/ENRICHMENTS_invitrodb_v3_1_BURST_hitc_MC_2019.tsv")
# extract aeid from 'Assay Name' column in aeid_table
aeid_table$aeid <- as.character(aeid_table$aeid)
library(stringr)
# make new column aeid using str_match
baseline$aeid <- str_match(baseline$`Assay Name`, ":\\s*(.*?)\\s*_")[,2]
burst$aeid <- str_match(burst$`Assay Name`, ":\\s*(.*?)\\s*_")[,2]

# match aenm from aeid_table by aeid
baseline_aenm <- merge( x = baseline, 
                        y = aeid_table[, c("aeid", "assay_component_endpoint_name") ],
                        by = "aeid", all.x = T)
burst_aenm <- merge( x = burst, 
                        y = aeid_table[, c("aeid", "assay_component_endpoint_name") ],
                        by = "aeid", all.x = T)

write.csv(burst_aenm, "OutputData/burst_aenm.csv")
write.csv(baseline_aenm, "OutputData/baseline_aenm.csv")


```

Visualizing the data. Using "label" of ToxPrint and "assay_component_endpoint_name" as the units of analysis for  independent categorical variables. Dependent variables will be
annotation attribute 

Q1: What is the count distribution of label?
Investigating prevelance in assay_component_endpoint_name [prmoiscuity]
Q2: What is the count distribution of assay_component_endpoint_name?
Investigating levels of diversity. [diversity]

Data preparation:
collapse label by assay_component_endpoint_name
collapse assay_component_endpoint_name by label 
```{r}
baseline_table <- function(baseline){
baseline_label_table<- table(baseline$label)
baseline_txp_table<- table(baseline$descriptors_name)

baseline_aeid_table<- table(baseline$aeid)
# for txp label
baseline_label_table_df <- as.data.frame(baseline_label_table)
names(baseline_label_table_df) <- c('ToxPrint_label', 'enriched assay count')
# for txp number
baseline_txp_table_df <- as.data.frame(baseline_txp_table)
names(baseline_txp_table_df) <- c('ToxPrint', 'enriched assay count')
return(baseline_label_table_df)
}
```



Future: density plot can show overall distributio; perhaps label groups of ToxPrint on it to show most promiscuous and leas, etc.n

Try wordcloud
```{r}
# making instance of baseline table
baseline_label_table_df <- baseline_table(baseline)

```

```{r}
library(wordcloud)
dev.new(width = 10000, height = 10000, unit = "px")
set.seed(100)
wordcloud(baseline_label_table_df$ToxPrint_label, baseline_label_table_df$`enriched assay count`, min.freq=100, max.words = 10, ,  colors=brewer.pal(6, "Dark2"), scale=c(4,.5), )

```
trying wordcloud with ggplot2
```{r}
library(ggplot2)
library(ggwordcloud)
wordcloud_txp <- function(baseline_table,top, seed ){
baseline_label_table_df_top_10 <- baseline_label_table_df[order(-baseline_label_table_df$`enriched assay count`),][1:top,]
# plotting wordcloud for top 20 ToxPrint labels
set.seed(seed)
ggplot(baseline_label_table_df_top_10, 
       aes(label = baseline_label_table_df_top_10$ToxPrint_label, 
           size = baseline_label_table_df_top_10$`enriched assay count`,
           #color = factor(sample.int(10, nrow(baseline_label_table_df_top_10), replace = TRUE)) ) # color randomly assigned
            
         color = baseline_label_table_df_top_10$`enriched assay count`)) +
  geom_text_wordcloud() +
  #geom_text_wordcloud_area(shape = "diamond") +
  theme_minimal()
}
```
Making example images
```{r}
# for baseline
#tiff(file="OutputImages/baseline_top20_45.tiff")
wordcloud_txp(baseline_table(baseline), 20, 45)


```
```{r}
#tiff(file="OutputImages/baseline_top20_45.tiff")
# for burst
wordcloud_txp(baseline_table(burst), 20, 45)


```
Rephrasing research questions:
QQ1) Which ToxPrints are no longer enriched or newly enriched upon application of filter (burst)?
```{r}
## perform set operations on ToxPrints in either burst or baseline groups
baseline_txp <- unique(baseline$label)
burst_txp <- unique(burst$label)
# get union
baseline_burst_union <- union(burst_txp, baseline_txp)
# get intersection
baseline_burst_intersect <- intersect(burst_txp, baseline_txp)
# get diff between burst and baseline ( burst specific or gained)
burst_specific <- setdiff(burst_txp, baseline_txp)
# get diff between baseline and burst ( baseline specific or lost)
baseline_specific <- setdiff(baseline_txp, burst_txp)
# draw a set diagram

```
draw a venn diagram to visualize the counts
```{r}
# Load library
library(VennDiagram)
 
```

try again
```{r}

# area1
area2 <- length(c(burst_specific, baseline_burst_intersect))
# area2
area1 <- length(c(baseline_specific, baseline_burst_intersect))
#cross.area
cross.area <- length(baseline_burst_intersect)


# A more complicated diagram Demonstrating external area labels
venn.plot <- draw.pairwise.venn(
area1 = area1,
area2 = area2,
cross.area = cross.area,
category = c("First", "Second"),
fill = c("blue", "red"),
lty = "blank",
cex = 2,
cat.cex = 2,
cat.pos = c(285, 105),
cat.dist = 0.09,
cat.just = list(c(-1, -1), c(1, 1)),
ext.pos = 30,
ext.dist = -0.05,
ext.length = 0.85,
ext.line.lwd = 2,
ext.line.lty = "dashed"
)
grid.draw(venn.plot)
grid.newpage()


```
  try euler diagram
    
```{r}
  # Demonstrating an Euler diagram
venn.plot <- draw.pairwise.venn(
area1 = area1,
area2 = area2,
cross.area = cross.area,
category = c("First", "Second"),
cat.pos = c(0, 180),
euler.d = TRUE,
sep.dist = 0.03,
rotation.degree = 45
)
# Writing to file
tiff(filename = "OutputImages/burst_baseline_euler_diagram.tiff", compression = "lzw")
grid.draw(venn.plot)
  
```

```{r}
venn.plot <- draw.pairwise.venn(
area1 = area1,
area2 = area2,
cross.area = cross.area,
category = c("baseline", "burst"),
fill = c("blue", "red"),
lty = "blank",
cex = 2,
cat.cex = 2,
cat.pos = c(285, 105),
cat.dist = 0.09,
cat.just = list(c(-1, -1), c(1, 1)),
ext.pos = 30,
ext.dist = -0.05,
ext.length = 0.85,
ext.line.lwd = 2,
ext.line.lty = "dashed"
)
# Writing to file
#tiff(filename = "OutputImages/burst_baseline_set_diagram.tiff", compression = "lzw")
grid.draw(venn.plot)



```
For overlapping ToxPrints, compute burst-baseline/baseline
```{r}
baseline_table_txp <- baseline_table(baseline)
burst_table_txp <- baseline_table(burst)
# change enriched assay count column to burst or baseline
names(baseline_table_txp)[names(baseline_table_txp) == "enriched assay count"] <- "baseline_count"
names(burst_table_txp)[names(burst_table_txp) == "enriched assay count"] <- "burst_count"
# 
burst_baseline_count <- merge(x = baseline_table_txp, y = burst_table_txp, by = "ToxPrint_label",
                              all.x = T)
# make differential normalized count column
setDT(burst_baseline_count)
burst_baseline_count[, relative_change := (burst_count - baseline_count) / baseline_count]
burst_baseline_count <- burst_baseline_count[!is.na(relative_change)]
burst_baseline_count <- burst_baseline_count[!is.na(ToxPrint_label)]

#



```

plot horizontal bar

```{r}
options(repr.plot.width=8, repr.plot.height=3)
# order asc
burst_baseline_count_ordered <- burst_baseline_count[order(relative_change)][1:40]
       
ggplot(burst_baseline_count_ordered, aes(x = ToxPrint_label, y = relative_change, main="Car Distribution")) +
         geom_bar(stat = "identity") +
         coord_flip() + scale_y_continuous(name="relative_difference_to_baseline") +
  scale_x_discrete(name="Chemical class") +
theme(axis.text.x = element_text(face="bold", color="#008000",
                           size=8, angle=0),
          axis.text.y = element_text(face="bold", color="#008000",
                           size=8, angle=0))


```
Re-formatting for making a map to visualize the interactions between assays according to chemotype enrichment profile
```{r}
# make two instances of the baeline enrichments of just the aenm and ToxPrint label
baseline_dup <- baseline_aenm # make a copt of baseline
names(baseline_dup)[names(baseline_dup) == "assay_component_endpoint_name"] <- "aenm"
names(baseline_dup)[names(baseline_dup) == "label"] <- "txp_nm"
baseline_dup_red <- baseline_dup[, c("aenm", "txp_nm")] # isolate the cols
baseline_dup_red_dip <- baseline_dup_red # make another instance
names(baseline_dup_red)[names(baseline_dup_red) == "aenm"] <- "aenm1" # rename the aenm 
names(baseline_dup_red_dip)[names(baseline_dup_red_dip) == "aenm"] <- "aenm2" # rename the aenm 
# merge the duplicated tables
baseline_merged_by_txpnm <- merge( x = baseline_dup_red, y = baseline_dup_red_dip, 
                                   all.x = T, by = "txp_nm", allow.cartesian = T)
# remove the self links
baseline_merged_by_txpnm_unique <- baseline_merged_by_txpnm[aenm1!= aenm2]
write.csv(baseline_merged_by_txpnm_unique, "baseline_merged_by_txpnm_unique.csv")
write.csv(baseline_merged_by_txpnm_unique[1:1000], "baseline_merged_by_txpnm_unique_first_1000.csv")


```
make a enrichment profile per assay

```{r}
txp_fp <- fread("txp_fp.csv")
# make new col on the file
library(reshape2)
baseline_dup_red[, n:=1] # dummy variable for presence of feature
baseline_spread <- reshape2::dcast(baseline_dup_red, aenm1~txp_nm, value.var = "n", fill = 0)
# 


```
use fingerprint and base packages to calculate Tanimoto similarity and other fingerprint stuff

```{r}
# inititalize empty matrix
aenm_mat <- matrix(data = NA, nrow = nrow(baseline_spread), ncol = nrow(baseline_spread), byrow = FALSE,
            dimnames = list(unique(baseline_spread$aenm1)))

colnames(aenm_mat) <- unique(baseline_spread$aenm1)
# setup for loop
for(i in 1:nrow(baseline_spread)){
  for(j in 1:nrow(baseline_spread)){
  
  aenm_mat[i,j] = jaccard(baseline_spread[i, 2:461], baseline_spread[j, 2:461])
  
  }
}

```